---
title: "Data Dictionary - Emory QTM Fall 2022"
output: html_document
author: Tony Chen Gu, Nancy Chen
date: '2022-11-14'
---
## Tony - 2022 Fall
```{r}
data = read.csv("export_dataframe_stage2.csv")
```

### air_tmep
- record the temperature in Celsius
- 60 percent of data is not NA and 0.5% of them shows a temperature over 40 degrees Celsius (104 F)
```{r}
summary(data$air_temp)
num_obs = length(data$air_temp)
num_NA = sum(is.na(data$air_temp))
paste("number of NA is ",num_NA)

paste("numer of valid data is ",num_obs - num_NA)
paste("percent of NA is",num_NA/ num_obs)
paste("percent of valid data is",(num_obs - num_NA) / num_obs)
```
Min to 3rd Quantile looks good, the Max value looks weird.
```{r}
over40 = sum(data$air_temp > 40, na.rm = TRUE)
paste("There are", over40, "of recording shows a temperature over 40 degrees Celcius (104 F). It's", over40/(num_obs - num_NA), "% of valid data.")
```
### air_tmeperature
- record the temperature in Celsius
- shows more NA than air_temp (60% of data is NA)
- 0.3% of valid data is higher than 40 degrees Celsius (104 F)
```{r}
summary(data$air_temperature)
num_obs = length(data$air_temperature)
num_NA = sum(is.na(data$air_temperature))
paste("number of NA is ",num_NA)

paste("numer of valid data is ",num_obs - num_NA)
paste("percent of NA is",num_NA/ num_obs)
paste("percent of valid data is",(num_obs - num_NA) / num_obs)
```
Min to 3rd Quantile looks good, the Max value looks weird.
```{r}
over40 = sum(data$air_temperature > 40, na.rm = TRUE)
paste("There are", over40, "of recording shows a temperature over 40 degrees Celcius (104 F). It's", over40/(num_obs - num_NA), "% of valid data.")
```

### min_temp
- too much NAs (80%) makes it almost unusable
- only 3% of the valid data is higher than 35 degree Celcius (95 F)
```{r}
summary(data$min_temp)
num_obs = length(data$min_temp)
num_NA = sum(is.na(data$min_temp))
paste("number of NA is ",num_NA)

paste("numer of valid data is ",num_obs - num_NA)
paste("percent of NA is",num_NA/ num_obs)
paste("percent of valid data is",(num_obs - num_NA) / num_obs)
```
min_temp of 95 is defiantly abnormal
```{r}
over35 = sum(data$air_temperature > 35, na.rm = TRUE)
paste("There are", over35, "of recording shows a temperature over 35 degree Celcius (95 F). It's", over35/(num_obs - num_NA), "% of valid data.")
```

### max temp
- unsure if it's about air or water temperature
- only about 10% of valid data
- has 0.3% of data over 40 degree
```{r}
summary(data$max_temp)
num_obs = length(data$max_temp)
num_NA = sum(is.na(data$max_temp))
paste("number of NA is ",num_NA)

paste("numer of valid data is ",num_obs - num_NA)
paste("percent of NA is",num_NA/ num_obs)
paste("percent of valid data is",(num_obs - num_NA) / num_obs)
```
Max_temp over 355 is weird.
```{r}
over40 = sum(data$air_temperature > 40, na.rm = TRUE)
paste("There are", over40, "of recording shows a temperature over 40 degrees Celcius (104 F). It's", over40/(num_obs - num_NA), "% of valid data.")
```

### water_temp
- record the water temperature in Celsius
- about 40% of the data are NA
- has abnormal data like 110 degree, only 0.3% of data is over 40 degree
```{r}
summary(data$water_temp)
num_obs = length(data$water_temp)
num_NA = sum(is.na(data$water_temp))
paste("number of NA is ",num_NA)

paste("numer of valid data is ",num_obs - num_NA)
paste("percent of NA is",num_NA/ num_obs)
paste("percent of valid data is",(num_obs - num_NA) / num_obs)
```
water temperature over 100 degree is too high. Look more into its data.
```{r}
over40 = sum(data$water_temp > 40, na.rm = TRUE)
paste("There are", over40, "of recording shows a water temperature over 40 degrees Celcius (104 F). It's", over40/(num_obs - num_NA), "% of valid data.")
hist(data$water_temp)
```
  
## Nancy - 2022 Fall
import dataset: 
```{r}
stage2<-read.csv("export_dataframe_stage2.csv")
```

### stream 
status of the stream 

- 3 distinct status: `rocky`, `muddy`, `muddy;rocky`
- All strings 
- 197 empty values, 5 numeric values (most likely to be input errors)

```{r}
# display unique values
unique(stage2$stream)

# counting for the number of null/error values
count_empty = 0
count_error = 0
for (i in 1:length(stage2$stream)){
  if (stage2$stream[i] == ""){
    count_empty = count_empty +1
  }
  if (stage2$stream[i] == "27" |stage2$stream[i] == "26.67" | stage2$stream[i] == "23" |stage2$stream[i] == "8.8" | stage2$stream[i] == "10.5" ){
    count_error = count_error + 1
  }
}
  
paste0("count_empty: ", count_empty) 
paste0("count_error: ", count_error) 
```

### method 
method used to measure the values

* 2 distinct method: `kick`, `d-frame`
  + **kick**: Measuring the abundance of macroinvertebrates (small aquatic animals) in the river by kicking up gravels. Unpolluted river will have a higher diversity  
  + **d-frame**: sweeping or jabbing nets into appropriate habitats to collect organisms from the sediment and water column
* All strings 
* 59325 empty values, 4 numeric values (most likely to be input errors)

```{r}
# display unique values
unique(stage2$method)

# counting null/error values
count_empty = 0
count_error = 0
for (i in 1:length(stage2$method)){
  if (stage2$method[i] == ""){
    count_empty = count_empty +1
  }
  if (stage2$method[i] == "22.2" |stage2$method[i] == "21.11" | stage2$method[i] == "22.6" |stage2$method[i] == "12.7"){
    count_error = count_error + 1
  }
}

#printing
paste0("count_empty: ", count_empty) 
paste0("count_error: ", count_error) 
```

### wqi
**Water quality limited segment**: any stream segment where the water quality does not or will not meet applicable water quality standards

- 34 distinct values 
- 59084 NA values 
- All integers 

```{r}
# display unique values
unique(stage2$wqi)

# counting null values
count_NA = 0
for (i in 1:length(stage2$wqi)){
  if (is.na(stage2$wqi[i])){
    count_NA = count_NA +1
  }
}

# printing
paste0("count_NA: ", count_NA) 
```

### habitat
natural environment around the stream

- 48 distinct habitats
- All strings 
- 59144 empty values

```{r}
# display a few unique values
head(unique(stage2$habitat))

# count the number of distinct values 
length(unique(stage2$habitat))

# count empty values 
count_empty = 0
for (i in 1:length(stage2$habitat)){
  if (stage2$habitat[i] == ""){
    count_empty = count_empty +1
  }
}

# printing
paste0("count_empty: ", count_empty) 
```


### chem_detail_rid 
chemical details of the stream 

- 58728 distinct values 
- All numeric integers
- No null values

```{r}
# display a few distinct values 
head(unique(stage2$chem_detail_rid))

# number of distinct values
length(unique(stage2$chem_detail_rid))

# counting number of empty values (check all possible way of null values inputs)
count_empty = 0
for (i in 1:length(stage2$chem_detail_rid)){
  if (stage2$chem_detail_rid[i] == "NULL" | stage2$chem_detail_rid[i] == "" | is.na(stage2$chem_detail_rid[i])){
    count_empty = count_empty +1
  }
}

# printing
paste0("count_empty: ", count_empty) 
```